Demystifying the Computer Science \\ PhD Admission in the US\\{\large A Handbook for International \emph{and Domestic} Students

### 2024-10-30 demystify.tex
updates

Even if you do everything right, you will likely get a generic (\autoref{sec:interpreting-response}) or no response from a prof. Don't take it personally. They might not have the bandwidth to take on new students (very common), or they might not be interested in your background.

\subsection{Interpreting Response}\label{sec:interpreting-response}

Some generic responses are:
\begin{itemize}
\item \textbf{No reply}. Very common. You might try again in a few weeks or months, but don't expect a reply. And after a couple of tries, you should move on.  It simply means they are not interested.
\item \textbf{Not taking students but encourage you to apply}. Polite way of saying not interested and referring you to the admission process.
\item \textbf{Not taken student this year (but encourage you to apply next year)}. Polite but generic response.  But If you do apply next year, you can mention that you contacted them last year and they encouraged you to apply.
\item \textbf{Come talk to me after you're admitted}. Generic. Refer you to the admission process. But if you get admitted then you can reply to them and say you're admitted and would like to talk.
\item \textbf{Cannot admit you directly. Need to go through admission process}. Generic. They are not interested and refer you to the admission process.
\item \textbf{I am not taking students but I co-advise/can serve on your committee} While this might sound good, it's generic because it says once you're admitted and have an adviser, then contact me again. 
\item \textbf{I am not taking student but Prof. X might be}. Not common practice as most profs. will not refer you to their colleagues. However, this is better than the previous responses. While they cannot take you, they think you are a good fit for X. So follow up with a thank you and say you'll contact X.  And then contact X and say that Y referred you to them.
\end{itemize}

@@

@@

\section{Are profs. very busy (and therefore ghost me)?}\label{sec:busy}

Yes, profs. are busy. They have many deadlines, meetings, and emails, many of which are from prospective students looking for research opportunities.  Tenure-line profs. (\autoref{sec:faculty-types}) typically have 5 jobs (research, teaching, service, grant writing, and mentoring students). They also have a life outside of work, e.g., family, hobbies, etc.  So yes, they are super busy.

However, this is not why they ``ghost'' you. The main reason is that they are not interested. They might not have funding, the bandwidth to take on new students, or they might not be interested in your background. So don't take it personally if you don't get a reply.  If you want to increase your chance of getting a reply, see \autoref{sec:contact}.

\begin{commentbox}{Vu}
I scan through \emph{every single email} sent by students for research opportunities (many of which are from undergrads and highschool students). However, I only reply to those that I think are a good fit and ask them to chat with me. I tend to be more responsive to undergrads and highschool students , especially those in the local area, to encourage them to pursue research.  
\end{commentbox}

>>>>

\subsection{Interpreting Response}\label{sec:interpreting-response}
Even if you do everything right, you will likely get a generic (\autoref{sec:interpreting-response}) or no response from a prof. Don't take it personally. They might not have the bandwidth to take on new students (very common), or they might not be interested in your background.

Some generic responses are:
\begin{itemize}
\item \textbf{No reply}. This is by far the most common response (see why in \autoref{sec:busy}).  It means they are not interested.
You might try again in a few weeks or months, but don't expect a reply. And after a couple of tries, you should move on.  It simply means they are not interested.
\item \textbf{Not taking students but encourage you to apply}. Polite way of saying not interested and referring you to the admission process. Note that this does not in any way mean that they think you have a good chance of getting admitted.

\item \textbf{Not taken student this year (but encourage you to apply next year)}. Polite but generic response.  And like the previous one, encouragement to apply does not mean they think you have a good chance of getting admitted.

\item \textbf{Come talk to me after you're admitted}. Generic. Refer you to the admission process. But if you get admitted then you can reply to them and say you're admitted and would like to talk.

\item \textbf{Cannot admit you directly. Need to go through admission process}. Generic. They are not interested and refer you to the admission process.

\item \textbf{I am not taking students but I co-advise/can serve on your committee} While this might sound good, it's generic because it says once you're admitted and have an adviser, then contact me again.

\item \textbf{I am not taking student but Prof. X might be}. Not common as most profs. will not refer you to their colleagues. However, this is better than the previous responses. While they cannot take you, they think you are a good fit for X. So follow up with a thank you and say you'll contact X.  And then contact X and say that Y referred you to them.
\end{itemize}

@@

\section{Are profs. so busy that they completely ignore emails?} \label{sec:busy}

Profs. are busy. They have many deadlines, meetings, and emails, many of which are from prospective students looking for research opportunities. They also have a life outside of work, e.g., family, hobbies, etc.

But this is not why they ``ghost'' you and provide no response.  The main reason is that a response, no matter how generic, might be misinterpreted by the student.  For example, if they say they are not taking students but still (strongly) encourage you to apply, the student might take it as a positive sign of admission. Or if they say they are not interested, the student might take it as a personal rejection.

So it is safer to not reply at all.  This is why you should not take it personally if you do not get a reply.  It is not because profs. are busy, but because they are trying to avoid misunderstandings.

\begin{commentbox}{Vu: }
    I do scan through \emph{every single email} sent by students for admission and research opportunities (many of which are from undergrads and highschool students). However, I only reply to those that I think are a good fit and ask them to chat with me. 
    \end{commentbox}

@@
